---
title: "Analysis For One Lake"
author: " "
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=4.5)

library(ggplot2)
library(dplyr)
library(lubridate)

options(show.signif.stars=FALSE)
```



This document goes through the analysis for assessing trend over time in the maximum total cell density for blue-green algae for one reservoir.  This same procedure can be repeated for other reservoirs, and the exploratory plots should be looked at in each case to be sure the analysis looks reasonable.  

This analysis assumes the following:

1. The data available are appropriate for looking at a trend in the maximum summer value over years available

2. We are interested in the linear component of the trend over the available years, on the natural log scale.  Therefore, we are assuming explonential growth on the original scale.  Investigating the linear component over the entire time series does not assume there is no interesting behavior in shorter intervals within the time series. 

3. Will add more here

```{r loaddata, include=FALSE}
load("BGtotals_maxsummer.Rdata")
names(BGtotals_maxsummer)

load("BGtotals_data.Rdata")
names(BGtotals_data)
```


```{r chooselake}
l <- "MSR"
```


# Plots of time series

These plots include all observations available (black open circles), observations from June 1- Sept. 30 are colored in orange, and the maximum for each summer season (one for each year) is colored in magenta.  The EPA health guidelines are the horizontal lines.

```{r PlotMax_alldata}
  plot.dat <- filter(BGtotals_data, lake==l, depth_cat=="0-3 ft")
  plot.dat2 <- filter(BGtotals_data, month(date1) %in% c(5,6,7,8), lake==l, depth_cat=="0-3 ft")
  plot.dat3 <- filter(BGtotals_maxsummer, lake==l)
  #View(plot.dat2)
  
   ggplot() + 
        geom_hline(aes(yintercept=23.03), colour="red", size=0.5)+
        geom_hline(aes(yintercept=18.42), colour="pink", size=0.5) +
        geom_hline(aes(yintercept=16.81), colour="orange", size=0.5) +
        geom_point(data=plot.dat, aes(x=date1, y=log(tot_cells_Rmax)), col=1, shape=1, size=2) +
        #geom_line(data=plot.dat, aes(x=date1, y=log(tot_cells_Rmax))) + 
        geom_point(data=plot.dat2, aes(x=date1, y=log(tot_cells_Rmax)), col="orange", shape=16,
                   size=1) +
        #geom_line(data=plot.dat2, aes(x=date1, y=log(tot_cells_Rmax))) + 
        geom_point(data=plot.dat3, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax)),
                    col="magenta", shape=10) +
        geom_line(data=plot.dat3, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax)),
                    col=gray(0.3)) + 
        ggtitle(paste(l, "Maximum Summer Total Cells/L BG")) +
        scale_x_date(date_minor_breaks = "1 year") +
        ylim(5,25) +
        ylab("ln(Total cells/L)") + xlab("Sampling date")
```


```{r PlotMax2}
  plot.dat3 <- filter(BGtotals_maxsummer, lake==l, year!=1987)

   ggplot() + 
        geom_hline(aes(yintercept=23.03), colour="red", size=0.5)+
        geom_hline(aes(yintercept=18.42), colour="pink", size=0.5) +
        geom_hline(aes(yintercept=16.81), colour="orange", size=0.5) +
        geom_point(data=plot.dat3, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax), size=num_sampdates_per_max), col="magenta", alpha=0.7) +
        geom_line(data=plot.dat3, aes(x=datemax_cells_Ravg, y=log(maxtot_cells_Rmax)),
                    col=gray(0.3)) + 
        ggtitle(paste(l, "Maximum Summer Total Cells/L BG")) +
        scale_size_continuous("# Dates")+
        scale_x_date(date_minor_breaks = "1 year") +
        ylim(5,25) +
        ylab("ln(Total cells/L)") + xlab("Sampling date")

   ggplot() + 
        geom_hline(aes(yintercept=23.03), colour="red", size=0.5)+
        geom_hline(aes(yintercept=18.42), colour="pink", size=0.5) +
        geom_hline(aes(yintercept=16.81), colour="orange", size=0.5) +
        geom_point(data=plot.dat3, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax), size=num_sampdates_per_max, colour=as.factor(month(datemax_cells_Rmax))), alpha=0.7) +
        geom_line(data=plot.dat3, aes(x=datemax_cells_Ravg, y=log(maxtot_cells_Rmax)),
                    col=gray(0.3)) + 
        ggtitle(paste(l, "Maximum Summer Total Cells/L BG")) +
        scale_size_continuous("# Dates")+
        scale_colour_discrete("Month")+
        scale_x_date(date_minor_breaks = "1 year") +
        ylim(5,25) +
        ylab("ln(Total cells/L)") + xlab("Sampling date")
```


Concerns based on the plots for EFR:
- 1995 (one low value to choose from), and 2007 (one sampling occasions), though others also have only one occasion and are higher.
- Still issues with 2013 -- should really be 12 sampling dates I don't think.



# Estimate the linear trend

The linear component of the trend, as estimated by ordinary least squares is shown in the below plot.  The ribbon shows point-wise confidence intervals, not confidence bands around the whole the regression line.

```{r PlotMax}
  plot.dat3 <- filter(BGtotals_maxsummer, lake==l, year!=1987)

   ggplot() + 
        geom_hline(aes(yintercept=23.03), colour="red", size=0.5)+
        geom_hline(aes(yintercept=18.42), colour="pink", size=0.5) +
        geom_hline(aes(yintercept=16.81), colour="orange", size=0.5) +
        geom_point(data=plot.dat3, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax), size=num_sampdates_per_max),
                    col="magenta", alpha=0.7) +
        geom_line(data=plot.dat3, aes(x=datemax_cells_Ravg, y=log(maxtot_cells_Rmax)),
                    col=gray(0.3)) + 
        geom_smooth(data=plot.dat3, aes(x=datemax_cells_Ravg, y=log(maxtot_cells_Rmax)), method="lm", col="magenta")+
        ggtitle(paste(l, "Maximum Summer Total Cells/L BG")) +
        scale_size_continuous("# Dates")+
        scale_x_date(date_minor_breaks = "1 year") +
        ylim(5,25) +
        ylab("ln(Total cells/L)") + xlab("Sampling date")
   
  # geom_smooth(data=plot.dat3, aes(x=datemax_cells_Ravg, y=log(maxtot_cells_Rmax)), method = "lm", formula = y ~ splines::bs(x, 3), se = FALSE)+
```

```{r OLS, include=FALSE}
lm_out <- lm(log(maxtot_cells_Rmax) ~ datemax_cells_Rmax, data=filter(BGtotals_maxsummer, lake==l, year!=1987))

summary(lm_out)

#make a new X matrix to get predictions at
newx <- data.frame(datemax_cells_Rmax=rep(seq(min(na.exclude(BGtotals_maxsummer$datemax_cells_Rmax)), max(na.exclude(BGtotals_maxsummer$datemax_cells_Rmax)), length=100)))
lmpred_pwCIs <- data.frame(predict(lm_out, newdata=newx, interval="confidence", level=0.95))
lmpred_pwCIs_plotdat <- data.frame(datemax_cells_Rmax=newx[,1], fit=lmpred_pwCIs[,1], lcl=lmpred_pwCIs[,2], ucl=lmpred_pwCIs[,3])
```

```{r OLSweighted, include=FALSE}
lm_out_wtd <- lm(log(maxtot_cells_Rmax) ~ datemax_cells_Rmax, weights=num_sampdates_per_max, data=filter(BGtotals_maxsummer, lake==l, year!=1987))
summary(lm_out_wtd)

#make a new X matrix to get predictions at
lmpred_wtd_pwCIs <- data.frame(predict(lm_out_wtd, newdata=newx, interval="confidence", level=0.95))
lmpred_wtd_pwCIs_plotdat <- data.frame(datemax_cells_Rmax=newx[,1], fit=lmpred_wtd_pwCIs[,1], lcl=lmpred_wtd_pwCIs[,2], ucl=lmpred_wtd_pwCIs[,3])
```

In the following plots blue is ordinary least squares.  The orange is weighted least squares with weights equal to the number of sampling occasions for that year (not the number of stations, just the number of dates).  So, when there are more sampling dates available to obtain the maximum over, that observation is weighted more heavily in the linear trend estimation. [This is just a preliminary look at weighting the different maximums differently based on data used to obtain them.]

The ribbons are 95% point-wise confidence intervals.  I will plot confidence bands for the regression line later, which will be a little wider and are meant to cover the entire regression line simultaneously, not just the mean at a particular time (x value).

```{r plotWeighted}
#names(BGtotals_maxsummer)
plot.dat <- filter(BGtotals_maxsummer, lake==l, year!=1987)
#dim(plot.dat)

# Just lines
ggplot(plot.dat) + 
       geom_point(aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax), size=num_sampdates_per_max),
                    col="magenta", alpha=0.7) +
       geom_line(data=lmpred_pwCIs_plotdat, aes(y=fit, x=datemax_cells_Rmax), colour="blue", size=1)+
       geom_line(data=lmpred_wtd_pwCIs_plotdat, aes(y=fit, x=datemax_cells_Rmax), colour="orange", size=1)+
        ggtitle(paste(l, "Maximum Summer Total Cells/L BG")) +
        scale_size_continuous("# Dates")+
        scale_x_date(date_minor_breaks = "1 year") +
        ylim(5,25) +
        ylab("ln(Total cells/L)") + xlab("Sampling date")

# With ribbons
ggplot(plot.dat) + 
       geom_point(aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax), size=num_sampdates_per_max),
                    col="magenta", alpha=0.7) +
       geom_line(data=lmpred_pwCIs_plotdat, aes(y=fit, x=datemax_cells_Rmax), colour="blue", size=1)+
       geom_line(data=lmpred_wtd_pwCIs_plotdat, aes(y=fit, x=datemax_cells_Rmax), colour="orange", size=1)+
       geom_ribbon(mapping = aes(ymax = ucl, ymin = lcl, x=datemax_cells_Rmax), data = lmpred_pwCIs_plotdat, alpha = 0.4, fill = "blue") +
       geom_ribbon(mapping = aes(ymax = ucl, ymin = lcl, x=datemax_cells_Rmax), data = lmpred_wtd_pwCIs_plotdat, alpha = 0.4, fill = "orange") +
        ggtitle(paste(l, "Maximum Summer Total Cells/L BG")) +
        scale_size_continuous("# Dates")+
        scale_x_date(date_minor_breaks = "1 year") +
        ylim(5,25) +
        ylab("ln(Total cells/L)") + xlab("Sampling date")

```


```{r PlotBacktransformed}
#make a new X matrix to get predictions at
newx <- data.frame(datemax_cells_Rmax=rep(seq(min(plot.dat$datemax_cells_Rmax), max(plot.dat$datemax_cells_Rmax), length=100)))

lmpred_pwCIs_plotdat_bt <- data.frame(datemax_cells_Rmax=newx[,1], fit=exp(lmpred_pwCIs[,1]), lcl=exp(lmpred_pwCIs[,2]), ucl=exp(lmpred_pwCIs[,3]))

lmpred_wtd_pwCIs_plotdat_bt <- data.frame(datemax_cells_Rmax=newx[,1], fit=exp(lmpred_wtd_pwCIs[,1]), lcl=exp(lmpred_wtd_pwCIs[,2]), ucl=exp(lmpred_wtd_pwCIs[,3]))


ggplot() + 
       geom_point(data=plot.dat, aes(x=datemax_cells_Rmax, y=maxtot_cells_Rmax, size=num_sampdates_per_max), col="magenta", alpha=0.7) +
       geom_line(data=lmpred_pwCIs_plotdat_bt, aes(y=fit, x=datemax_cells_Rmax), colour="blue", size=1)+
       geom_ribbon(mapping = aes(ymax = ucl, ymin = lcl, x=datemax_cells_Rmax), data = lmpred_pwCIs_plotdat_bt, alpha = 0.4, fill = "blue") +
       ggtitle(paste(l, "Maximum Summer Total Cells/L BG")) +
        scale_size_continuous("# Dates")+
        scale_x_date(date_minor_breaks = "1 year") +
        ylab("Total cells/L") + xlab("Sampling date")


ggplot() + 
       geom_point(data=plot.dat, aes(x=datemax_cells_Rmax, y=maxtot_cells_Rmax, size=num_sampdates_per_max), col="magenta", alpha=0.7) +
       geom_line(data=lmpred_pwCIs_plotdat_bt, aes(y=fit, x=datemax_cells_Rmax), colour="blue", size=1)+
       geom_ribbon(mapping = aes(ymax = ucl, ymin = lcl, x=datemax_cells_Rmax), data = lmpred_pwCIs_plotdat_bt, alpha = 0.4, fill = "blue") +
       geom_line(data=lmpred_wtd_pwCIs_plotdat_bt, aes(y=fit, x=datemax_cells_Rmax), colour="orange", size=1)+
       geom_ribbon(mapping = aes(ymax = ucl, ymin = lcl, x=datemax_cells_Rmax), data = lmpred_wtd_pwCIs_plotdat_bt, alpha = 0.4, fill = "orange") +
       ggtitle(paste(l, "Maximum Summer Total Cells/L BG")) +
        scale_size_continuous("# Dates")+
        scale_x_date(date_minor_breaks = "1 year") +
        ylab("Total cells/L") + xlab("Sampling date")

```


Plot the confidence bands, rather than point-wise confidence intervals.

```{r PlotBandsNotPointwise}
#
```


# Look at residuals


Here, we take a look at the residuals from both the OLS fit and the weighted fit.

```{r residplot, fig.width=4, fig.height=3}
r_lm <- resid(lm_out)
 plot(fitted(lm_out), r_lm)
 abline(h=0)
 qqnorm(r_lm); qqline(r_lm)

r_wtd <- resid(lm_out_wtd)
 plot(fitted(lm_out_wtd), r_wtd)
 abline(h=0)
 qqnorm(r_wtd); qqline(r_wtd)
```


There is no evidence of serial autocorrelation after looking at the maximums.
```{r acfs, fig.width=4, fig.height=3}
acf(r_lm)
pacf(r_lm)

acf(r_wtd)
pacf(r_wtd)
```



