---
title: "Change Point Analysis"
author: "Megan Higgs"
date: "September 13, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bcp)
library(strucchange)
library(segmented)

#ADD path to working directory here
```

Here is some code for getting started with change point analysis.  This is just meant to get you started and is not a complete analysis.  I'm sure there are other references/methods/packages out there now as well.

```{r loaddata, include=FALSE}
load("BGtotals_maxsummer.Rdata")
names(BGtotals_maxsummer)

load("BGtotals_data.Rdata")
names(BGtotals_data)

#make a year column that is an integer
BGtotals_maxsummer <- mutate(BGtotals_maxsummer, year_int=year-1987)

#subset to lakes if you would like
lake_list <- c("EFR", "CHL", "RRR")
BGdata <- filter(BGtotals_maxsummer, lake %in% lake_list, year!=1987)
```

```{r breakpoint}
#Breakpoints function
 help(breakpoints) #See example code at end of help file

  break1.out <- breakpoints(log(maxtot_cells_Rmax+1) ~ datemax_cells_Rmax, h=0.15, breaks=1, het.reg=TRUE, het.err=TRUE, data=filter(BGdata, lake=="EFR"))
  summary(break1.out)
  plot(break1.out)
  
  filter(BGdata, lake=="EFR")$datemax_cells_Rmax[13] ##2007-06-20
  
#See help file code example for plotting the fitted lines?  
  
```  
  

```{r segmented} 
  ###### Fit a broken-line model using the segmented() package in R - constrains the lines to meet
   ####  at the changepoint

#NEED TO CHANGE 
    lin.out <- lm(log(maxtot_cells_Rmax+1) ~ year_int, data=filter(BGdata, lake=="EFR")) 
     seg.lin.out <- segmented(lin.out, seg.Z=~year_int, psi=13, control=seg.control(it.max=100))
     summary(seg.lin.out)
     slope.lin <- slope(seg.lin.out)
     slope.lin$year_int[,-(2:3)]
     exp(slope.lin$year_int[,-(2:3)]) 
     davies.test(lin.out, "year_int", k=5)  ###not working -- fix code
    
  
   #dev.new()
     #ISSUES WITH THE X-AXIS ON THIS PLOT relative to the use of time series object: ts()
   plot(BGdata$year_int[BGdata$lake=="EFR"], log(BGdata$maxtot_cells_Rmax[BGdata$lake=="EFR"]), type="b", pch=18, cex=1, main="",xlab="Year", xaxt="n") 
      #axis(1,at=seq(1,29,by=2), labels=as.character(1983:2011)[seq(1,29,by=2)])
      lines(ts(fitted(lin.out), start=1), col=2, lwd=2)
      lines(ts(fitted(seg.lin.out), start=1), col="purple", lwd=3)
```

NOTES:
-You can fit changepoint models that restrict the two segments to meet at the change point, or that allow a jump (completely separate line)
-Try not to put too much emphasis on selection of change point --- this is more useful as a describing differences around a suspected changepoint

