---
title: "Compare Measures"
author: " "
date: "September 27, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.with=7, fig.height=4, warning=FALSE, message=FALSE)
library(ggplot2)
library(plyr)
library(dplyr)
library(reshape2)
library(lubridate)
library(openxlsx)
#setwd("~/repos/EPA TO57/analysis") #SET TO YOUR WORKING DIRECTORY
```

In this document, we compare maximum summer cell density and biovolume for (1) Total across all taxa, (2) Total for cyanobacteria, (3) Total for toxin-producing. We also compare proportion cyanobacteria between cell density and biovlume. 

```{r loaddata}
#Load totals
load("totals_alltaxa_data.Rdata")
load("totals_alltaxaNONA_data.Rdata")
load("ALLtotals_maxsummer.Rdata")

#Load cyanobacteria totals
load("BGtotals_data.Rdata")
load("BGtotals_maxsummer.Rdata")

#Load toxin totals
load("toxin_totals.Rdata")
load("toxin_totals_maxsummer.Rdata")

#Load proportions cyanobacteria
load("BGdom_data.Rdata")
load("BGdom_data_NONA.Rdata")
load("BGdom_maxsummer.Rdata") #created using the NONA data set (0's in place of NAs)
```



#1. Comparing Summer Maximum Total Cell Density for all taxa, cyanobacteria, and toxin producing

```{r}
 lake_list <- unique(BGdom_maxsummer$lake)
 for (l in lake_list) {
  d <- ggplot() + 
        geom_point(data=filter(BGtotals_maxsummer, lake==l), aes(x=datemax_cells_Rmax, 
                      y=log(maxtot_cells_Rmax), size=num_sampdates_per_max, colour="cyan"), alpha=0.7) +
        geom_line(data=filter(BGtotals_maxsummer, lake==l), aes(x=datemax_cells_Ravg, y=log(maxtot_cells_Rmax)),
                    col=gray(0.3)) + 
       geom_point(data=filter(ALLtotals_maxsummer, lake==l), aes(x=datemax_cells_Rmax, 
                      y=log(maxtot_cells_Rmax), size=num_sampdates_per_max, colour=gray(0.2)), alpha=0.7) +
        geom_line(data=filter(ALLtotals_maxsummer, lake==l), aes(x=datemax_cells_Ravg, 
                      y=log(maxtot_cells_Rmax)), col=gray(0.3)) +
        geom_point(data=filter(toxin_totals_maxsummer, lake==l), aes(x=datemax_cells_Rmax, 
                      y=log(maxtot_cells_Rmax), size=num_sampdates_per_max, colour="red"), alpha=0.7) +
        geom_line(data=filter(toxin_totals_maxsummer, lake==l), aes(x=datemax_cells_Ravg,
                      y=log(maxtot_cells_Rmax)),col=gray(0.3)) +
        ggtitle(paste(l, "Maximum Summer Total Cell Density")) +
        scale_size_continuous("# Dates")+
        scale_x_date(date_minor_breaks = "1 year") +
        scale_colour_manual(" ", values=c(gray(0.2), "cyan", "red"), labels=c( "All taxa", "Cyanobacteria", "Toxin producing")) +
        ylab("ln(Total cells/L)") + xlab("Sampling date")
  print(d)
}
```


#2. Comparing Summer Maximum Total Biovolume for all taxa, cyanobacteria, and toxin producing

```{r}
 lake_list <- unique(BGdom_maxsummer$lake)
 for (l in lake_list) {
  d <- ggplot() + 
        geom_point(data=filter(BGtotals_maxsummer, lake==l), aes(x=datemax_cells_Rmax, 
                      y=log(maxtot_bv_Rmax), size=num_sampdates_per_max, colour="cyan"), alpha=0.7) +
        geom_line(data=filter(BGtotals_maxsummer, lake==l), aes(x=datemax_cells_Ravg, y=log(maxtot_bv_Rmax)),
                    col=gray(0.3)) + 
       geom_point(data=filter(ALLtotals_maxsummer, lake==l), aes(x=datemax_cells_Rmax, 
                      y=log(maxtot_bv_Rmax), size=num_sampdates_per_max, colour=gray(0.2)), alpha=0.7) +
        geom_line(data=filter(ALLtotals_maxsummer, lake==l), aes(x=datemax_cells_Ravg, y=log(maxtot_bv_Rmax)),
                    col=gray(0.3)) +
        geom_point(data=filter(toxin_totals_maxsummer, lake==l), aes(x=datemax_cells_Rmax, 
                      y=log(maxtot_bv_Rmax), size=num_sampdates_per_max, colour="red"), alpha=0.7) +
        geom_line(data=filter(toxin_totals_maxsummer, lake==l), aes(x=datemax_cells_Ravg, y=log(maxtot_bv_Rmax)),
                    col=gray(0.3)) +
        ggtitle(paste(l, "Maximum Summer Total Biovolume")) +
        scale_size_continuous("# Dates")+
        scale_x_date(date_minor_breaks = "1 year") +
        scale_colour_manual(" ", values=c(gray(0.2), "cyan", "red"), labels=c( "All taxa", "Cyanobacteria", "Toxin producing")) +
        ylab("ln(Total Biovolume)") + xlab("Sampling date")
  print(d)
}
```

#3. Compare maximum summer proportion of cyanobacteria between cell density and biovolume
```{r}
 lake_list <- unique(BGdom_maxsummer$lake)
 for (l in lake_list) {
  d <- ggplot() + 
        geom_point(data=filter(BGdom_maxsummer, lake==l), aes(x=datemax_cells_Rmax, 
                      y=maxrel_cells_Rmax, size=num_sampdates_per_max, colour="green"), alpha=0.6) +
        geom_line(data=filter(BGdom_maxsummer, lake==l), aes(x=datemax_cells_Ravg, y=maxrel_cells_Rmax),
                    col=gray(0.3)) + 
        geom_point(data=filter(BGdom_maxsummer, lake==l), aes(x=datemax_cells_Rmax, 
                      y=maxrel_bv_Rmax, size=num_sampdates_per_max, colour="blue"), alpha=0.6) +
        geom_line(data=filter(BGdom_maxsummer, lake==l), aes(x=datemax_cells_Ravg, y=maxrel_bv_Rmax),
                    col=gray(0.3)) +
        ggtitle(paste(l, "Maximum Summer Proportion Cyanobacteria")) +
        scale_size_continuous("# Dates")+
        scale_x_date(date_minor_breaks = "1 year") +
        scale_colour_manual(" ", values=c("green", "blue"), labels=c( "Biovolume", "Cell Density")) +
        ylim(0,1) +
        ylab("Proportion") + xlab("Sampling date")
  print(d)
 }
```
