---
title: "Exploratory Plots For One Lake"
author: " "
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
library(ggplot2)
library(plyr)
library(dplyr)
library(reshape2)
library(lubridate)
library(openxlsx)
setwd("~/repos/EPA TO57/analysis")
```

This document produces exploratory plots for blue-green algae in one reservoir.

```{r loadDatafiles}
load("BGtotals_data.Rdata") #Loads a data set named BGtotals_data
load("BGtotals_maxsummer.Rdata") #Loads a date set named BGtotals_maxsummer
```

```{r chooselake}
l <- "EFR"
```

#1. Plots of all data

```{r BGtotalcellsplot1, echo=FALSE}
#Panels for depth, colors by station
 #Original scale
  plot.dat <- filter(BGtotals_data, lake=="EFR", depth_cat %in% c("0-3 ft", "5 ft"))
  ggplot(plot.dat) + 
        geom_point(aes(x=date1, y=tot_cells_Ravg, colour=station)) +
        geom_line(aes(x=date1, y=tot_cells_Ravg, colour=station)) + 
        facet_wrap(~depth_cat, nrow=6, ncol=1) + 
        ggtitle(paste("EFR", "Blue-Green Algae")) +
        scale_x_date(date_minor_breaks = "1 year") +
        ylab("Total cells/L") +
        xlab("Sampling date")
  
 #Log scale
  ggplot(plot.dat, aes(x=date1, y=log(tot_cells_Ravg+1))) + 
        geom_point(aes(colour=station, shape=factor(ind_zero)), size=1.5) +
        geom_line(aes(colour=station)) + 
        facet_wrap(~depth_cat, nrow=2, ncol=1) + 
        ggtitle(paste(l, "Blue-Green Algae - log scale")) +
        scale_x_date(date_minor_breaks = "1 year") +
        ylab("log(Total cells/L)") +
        xlab("Sampling date") +
        ylim(c(-1,25)) + #ggplot does plot the log(0)'s at whatever the minimum is for the plot on the y-axis
        scale_shape_discrete(labels=c("count>0", "count=0")) +
        theme(legend.title=element_blank())
```

#2. Plots of Maximum summer counts relative to all data
```{r PlotMax_alldata}
  plot.dat <- filter(BGtotals_data, lake==l, depth_cat=="0-3 ft")
  plot.dat2 <- filter(BGtotals_data, month(date1) %in% c(5,6,7,8), lake==l, depth_cat=="0-3 ft")
  plot.dat3 <- filter(BGtotals_maxsummer, lake==l)
  #View(plot.dat2)
  
   ggplot() + 
        geom_hline(aes(yintercept=23.03), colour="red", size=0.5)+
        geom_hline(aes(yintercept=18.42), colour="pink", size=0.5) +
        geom_hline(aes(yintercept=16.81), colour="orange", size=0.5) +
        geom_point(data=plot.dat, aes(x=date1, y=log(tot_cells_Rmax)), col=1, shape=1, size=2) +
        #geom_line(data=plot.dat, aes(x=date1, y=log(tot_cells_Rmax))) + 
        geom_point(data=plot.dat2, aes(x=date1, y=log(tot_cells_Rmax)), col="orange", shape=16,
                   size=1) +
        #geom_line(data=plot.dat2, aes(x=date1, y=log(tot_cells_Rmax))) + 
        geom_point(data=plot.dat3, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax)),
                    col="magenta", shape=10) +
        geom_line(data=plot.dat3, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax)),
                    col=gray(0.3)) + 
        ggtitle(paste(l, "Maximum Summer Total Cells/L BG")) +
        scale_x_date(date_minor_breaks = "1 year") +
        ylim(5,25) +
        ylab("ln(Total cells/L)") + xlab("Sampling date")
```

#Just summer counts (June 1 - September 30)
```{r Only20001_JuneSept, echo=FALSE}
#load("BGtotals_data.Rdata")
lake_list <- c("EFR")
BGtotals_data <- dplyr::mutate(BGtotals_data, ind_June=ifelse(month(date1) %in% c(6,7,8,9), 0, 1))

for (l in lake_list) {
  plot.dat <- filter(BGtotals_data, lake==l, depth_cat %in% c("0-3 ft", "5 ft"),
                      station=="20001")
  d <- ggplot(plot.dat) + 
        geom_point(aes(x=date1, y=log(tot_cells_Ravg), colour=as.factor(ind_June)), size=2) +
        geom_line(aes(x=date1, y=log(tot_cells_Ravg))) + 
        facet_wrap(~depth_cat, nrow=6, ncol=1) + 
        ggtitle(paste(l, "20001- Blue-Green Algae - log scale")) +
        scale_x_date(date_minor_breaks = "1 year") +
        ylab("log(Total cells/L)") +
        xlab("Sampling date") +
        scale_color_discrete(labels=c("June-Aug", "Other")) +
        theme(legend.title=element_blank())
      print(d)
}
```

# Break up by months
```{r }
 plot.dat <- filter(BGtotals_data, lake==l, depth_cat %in% c("0-3 ft"), station %in% c("20001","20004","20024","21001"))
  d <- ggplot(plot.dat, aes(x=date1, y=log(tot_cells_Ravg))) + 
        geom_line(aes(colour=station)) + 
        geom_point(aes(colour=station), size=1.5) +
        facet_wrap(~month.f, nrow=5, ncol=2) + 
        ggtitle(paste(l, "Blue-Green Algae - MONTHS")) +
        scale_x_date(date_minor_breaks = "1 year") +
        ylab("log(Total cells/L)") +
        xlab("Sampling date")
      print(d)
```


## Panels for years

```{r}
day_in_yr <- as.Date(with(BGtotals_data, paste("1000","-",substring(date1,6,10), sep="")))
BGtotals_data$day_in_yr <- day_in_yr

BGtotals_maxsummer <- mutate(BGtotals_maxsummer, day_in_yr=as.Date(paste("1000","-",substring(datemax_cells_Rmax,6,10), sep="")),
year.f=factor(year))
```

```{r, fig.width=8, fig.height=6}
for (l in lake_list) {
plot.dat <- filter(BGtotals_data, lake==l, depth_cat=="0-3 ft", !year.f=="1995")
plot.dat2 <- filter(BGtotals_maxsummer, lake==l, !year==1995)
d5 <-  ggplot(plot.dat) +
  geom_point(data=plot.dat, aes(x=day_in_yr, y=log(tot_cells_Rmax), colour=station)) + 
        geom_vline(aes(xintercept=as.numeric(as.Date("1000-06-01"))), colour="red", size=0.5) +
        geom_vline(aes(xintercept=as.numeric(as.Date("1000-10-01"))), colour="red", size=0.5) +
        geom_point(data=plot.dat2, aes(y=log(maxtot_cells_Rmax), x=day_in_yr, group=year.f), colour="red", size=3) +
        geom_hline(aes(yintercept=23.03), colour="red", size=0.5)+
        geom_hline(aes(yintercept=18.42), colour="pink", size=0.5) +
        geom_hline(aes(yintercept=16.81), colour="orange", size=0.5) +
        facet_wrap(~year.f, nrow=5, ncol=4) +
        xlab("Day of year") +
        ylab("ln(BG Total density)")+
        ggtitle(paste(l)) +
        scale_x_date(date_minor_breaks = "1 month")
print(d5)
}
```


## Focus on the max's

```{r PlotMax2}
  plot.dat3 <- filter(BGtotals_maxsummer, lake==l, year!=1987)

   ggplot() + 
        geom_hline(aes(yintercept=23.03), colour="red", size=0.5)+
        geom_hline(aes(yintercept=18.42), colour="pink", size=0.5) +
        geom_hline(aes(yintercept=16.81), colour="orange", size=0.5) +
        geom_point(data=plot.dat3, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax), size=num_sampdates_per_max),
                    col="magenta", alpha=0.7) +
        geom_line(data=plot.dat3, aes(x=datemax_cells_Ravg, y=log(maxtot_cells_Rmax)),
                    col=gray(0.3)) + 
        ggtitle(paste(l, "Maximum Summer Total Cells/L BG")) +
        scale_size_continuous("# Dates")+
        scale_x_date(date_minor_breaks = "1 year") +
        ylim(5,25) +
        ylab("ln(Total cells/L)") + xlab("Sampling date")
```


Could add:

1. Plots from the original EFR analysis file

2. Plots from the Seasonality Investigation

3. Plots from the SummerMaxInvestigation 




