---
title: "Summer Max Investigation - Total Cell Density"
author: " "
date: "September 27, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.with=7, fig.height=4, warning=FALSE, message=FALSE)
library(ggplot2)
library(plyr)
library(dplyr)
library(reshape2)
library(lubridate)
library(openxlsx)
#setwd("~/repos/Phytoplankton-Data-Analysis/gitAnalysisFiles/analysis") #Change your working directory
```

Here is a look at using the maximum summer (June - Sept) blue-green algae total cell density as the response for each year compared to other counts.  


```{r loadData, eval=TRUE, include=FALSE}
load("BGtotals_data.Rdata") #Loads dataframe BGtotals_data
load("BGtotals_summer.Rdata") #Loads dataframe BGtotals_summer
load("BGtotals_maxsummer.Rdata")  #Loads dataframe BGtotals_maxsummer


## Best way to understand the plots is to run each chunck of plotting code just within R to
## see the output, rather than using Knit HTML to make the HTML file.  You can also set it to
## save all the plots in a folder when it compiles and usually I find this easiest using the
## Knit to Word (even if I don't care about having the actual word document).
```



#Comparing max to other counts

Legend info for below plots:  All counts for a lake (from all dates and all "good" stations) are shown in open black circles.  Orange points are samples taken in June 1 - September 30 (lumping all "good" stations).  The magenta points are the maximum value over all stations and sampling occasions each year between June 1 and September 30.  The orange, pink, and red horizontal lines correspond to the EPS guidelines for health risks: [EPA’s guidelines for health risks](https://www.epa.gov/nutrient-policy-data/guidelines-and-recommendations).
 
Rel. prob. acute health effects     |       Ln         |     Cells / L
----------------------------------- | ---------------- | ---------------------------------
             Low                    |       <16.81     |    <20,000,000      
           Moderate                 |    16.81-18.42   | 20,000,000 – 100,000,000     
            High                    |    18.42-23.03   | 100,000,000 – 10,000,000,000     
          Very high                 |       >23.03     |    >10,000,000,000  


```{r PlotsShowingMax, echo=FALSE, warning=FALSE, message=FALSE}
#lake_list <- c("EFR", "CHL", "RRR", "BHR")
lake_list <- unique(BGtotals_data$lake)
for (l in lake_list) {
  plot.dat <- filter(BGtotals_data, lake==l, depth_cat=="0-3 ft")
  plot.dat2 <- filter(BGtotals_summer, lake==l)
  plot.dat3 <- filter(BGtotals_maxsummer, lake==l)
  #View(plot.dat2)
  
  d <- ggplot() + 
        geom_point(data=plot.dat, aes(x=date1, y=log(tot_cells_Ravg)), col=1, shape=1, size=2) +
        #geom_line(data=plot.dat, aes(x=date1, y=log(tot_cells_Ravg))) + 
        geom_point(data=plot.dat2, aes(x=date1, y=log(tot_cells_Ravg)), col="orange", shape=16,
                   size=1) +
        #geom_line(data=plot.dat2, aes(x=date1, y=log(tot_cells_Ravg))) + 
        geom_point(data=plot.dat3, aes(x=datemax_cells_Ravg, y=log(maxtot_cells_Ravg)),
                    col="magenta", shape=10) +
        geom_line(data=plot.dat3, aes(x=datemax_cells_Ravg, y=log(maxtot_cells_Ravg)),
                    col=gray(0.3)) + 
        geom_hline(aes(yintercept=23.03), colour="red", size=0.5)+
        geom_hline(aes(yintercept=18.42), colour="pink", size=0.5) +
        geom_hline(aes(yintercept=16.81), colour="orange", size=0.5) +
        ggtitle(paste(l, "Blue-Green Algae - All good stations")) +
        scale_x_date(date_minor_breaks = "1 year") +
        ylim(5,25) +
        ylab("log(Total cells/L)") + xlab("Sampling date")
      print(d)
}
```



```{r, eval=FALSE, include=FALSE}
## Need to grab unique sampling dates for each year
BGtotals_data <- mutate(BGtotals_data, year=year(date1))

lake_list <- c("EFR", "CHL", "RRR", "BHR")
for (l in lake_list) {
  dat <- filter(BGtotals_data, lake==l, depth_cat=="0-3 ft")
  samp_dates <- with(dat, numeric(length(unique(date1))))
  year_list <- with(dat, sort(unique(year)))

  num_d_in_year <- c(0,with(dat, tapply(date1, year, function(x) length(unique(x)))))
  cumsum_d_in_year <- cumsum(num_d_in_year)
  
  for (j in 1:(length(year_list))) {
     samp_dates[(cumsum_d_in_year[j]+1):(cumsum_d_in_year[j+1])] <- with(filter(dat, year==year_list[j]), unique(yday(date1)))
  }
}
```


```{r, eval=FALSE, include=FALSE}
BGtotals_data <- mutate(BGtotals_data, year=year(date1))

lake_list <- c("EFR", "CHL", "RRR", "BHR")
for (l in lake_list) {
    num_d_in_year <- filter(BGtotals_data, lake==l, depth_cat=="0-3 ft") %>% 
                     dplyr::group_by(year) %>%     
                     dplyr::summarise(num_dates_in_year=length(unique(date1)))
    d <- ggplot() + geom_point(data=num_d_in_year, aes(y=num_dates_in_year, x=year))
    print(d)
}
```

## Day of sampling date vs. day of the year of the sampling

June 1 and September 30 are shown in orange horizontal lines.

```{r, eval=FALSE, include=FALSE}
## Generic version of day of year sampled over the years
ggplot() + 
  geom_point(data=filter(BGtotals_data, lake==l, depth_cat=="0-3 ft"), aes(x=date1, y=yday(date1),
                            colour=yday(date1))) +
             ylab("Day of Year (1-365)") +
             xlab("Date sampled") +
             scale_colour_gradient("Day of Year", limits=c(1,365), low="blue", high="red") +
             ylim(0,365) +
   geom_hline(aes(yintercept=c(154,275)), colour="orange") #June - September region
``` 


```{r}
## Color the date with the maximum count!  See above work to get this
lake_list <- with(BGtotals_data, unique(lake))
for (l in lake_list) {
plot.dat3 <- filter(BGtotals_maxsummer, lake==l)
d<- ggplot() + 
  geom_point(data=filter(BGtotals_data, lake==l, depth_cat=="0-3 ft"), aes(x=date1, y=yday(date1)), shape=1, size=2) +
  geom_point(data=plot.dat3, aes(x=datemax_cells_Rmax, y=yday(datemax_cells_Rmax)), col="red", shape=16) + 
  geom_line(data=plot.dat3, aes(x=datemax_cells_Rmax, y=yday(datemax_cells_Rmax)), col="red") +
             ylab("Day of Year (1-365)") +
             xlab("Date sampled") +
             ylim(0,365)+
   geom_hline(aes(yintercept=c(154,275)), colour="orange") +
   ggtitle(paste(l, " Timing of Max Cell Density - BG Algae")) +
   scale_x_date(date_minor_breaks = "1 month")
print(d)
}

 ##Number to choose from in each year IS increasing, BUT pattern does not look unusual -- issues with years with only one or two sampling occasion
```


# Look at different relationships between sampling dates, number of samples per year, and maximum log cell density

```{r AllSamples, eval=FALSE, include=FALSE}
#This code uses num_samps_per_max which includes number of sampling occasions (unique dates) AND the number of stations.  So, could look like we have a lot, but really it's just a lot of sampling locations (stations) at one sampling occasion, and what we're trying to look at is if there is a relationship between the maximum and the number of sampling dates in a particular year

for (l in lake_list) {

plot.dat3 <- filter(BGtotals_maxsummer, lake==l)

print(paste("LAKE = ", l))
 
d1 <- ggplot() + 
  geom_point(data=plot.dat3, aes(y=num_samps_per_max, x=datemax_cells_Rmax, size=log(maxtot_cells_Rmax), colour=as.factor(month(datemax_cells_Rmax))), alpha=0.5) + 
        ylab("Number Samples (dates & stations") + xlab("Sampling date") +
        scale_size_continuous("ln(Cell Dens)") +
        scale_colour_discrete("month") +
         ggtitle(paste(l)) +
        theme(legend.position="right")

ggplot() + 
  geom_point(data=plot.dat3, aes(y=num_sampdates_per_max, x=datemax_cells_Rmax, size=log(maxtot_cells_Rmax), colour=as.factor(month(datemax_cells_Rmax))), alpha=0.5) + 
        ylab("Number unique sampling dates") + xlab("Sampling date") +
        scale_size_continuous("ln(Cell Dens)") +
        scale_colour_discrete("month") +
         ggtitle(paste(l)) +
        theme(legend.position="right")


d2 <- ggplot() + 
    geom_point(data=plot.dat3, aes(x=num_samps_per_max, y=log(maxtot_cells_Rmax), colour=as.factor(month(datemax_cells_Rmax))), size=1.8) + 
          ylab("log(Total cells/L)") + xlab("Number of samples per year") +
          scale_colour_discrete("month") +
           ggtitle(paste(l)) +
          theme(legend.position="right")


d3 <- ggplot() + 
  geom_point(data=plot.dat3, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax), size=num_samps_per_max), colour="blue", alpha=0.6) + 
        ylab("log(Total cells/L)") + xlab("Sampling date") +
        scale_size_continuous("# Samples") +
         ggtitle(paste(l)) +
        theme(legend.position="right")

print(d1)
print(d2)
print(d3)

}
```



```{r}
for (l in lake_list) {
  
plot.dat3 <- filter(BGtotals_maxsummer, lake==l)

print(paste("LAKE = ", l))

d1 <- ggplot() + 
  geom_point(data=plot.dat3, aes(y=num_sampdates_per_max, x=datemax_cells_Rmax, size=log(maxtot_cells_Rmax), colour=as.factor(month(datemax_cells_Rmax))), alpha=0.5) + 
        ylab("Number unique sampling dates") + xlab("Sampling date") +
        scale_size_continuous("ln(Cell Dens)") +
        scale_colour_discrete("month") +
         ggtitle(paste(l)) +
        theme(legend.position="right")


  d2 <- ggplot() + 
    geom_point(data=plot.dat3, aes(x=num_sampdates_per_max, y=log(maxtot_cells_Rmax), colour=as.factor(month(datemax_cells_Rmax))), size=1.8) + 
          ylab("ln(Total cells/L)") + xlab("Number of sampling dates per year") +
          scale_colour_discrete("month") +
           ggtitle(paste(l)) +
          theme(legend.position="right")


d3 <- ggplot() + 
  geom_point(data=plot.dat3, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax), size=num_sampdates_per_max), colour="blue", alpha=0.6) + 
        ylab("ln(Total cells/L)") + xlab("Sampling date") +
        scale_size_continuous("# Dates") +
         ggtitle(paste(l)) +
        theme(legend.position="right")

print(d1)
print(d2)
print(d3)

}
```

# Look at all lakes lumped together for the relationships plotted above

```{r}
#All lakes together

ggplot() + 
  geom_point(data=BGtotals_maxsummer, aes(y=num_sampdates_per_max, x=datemax_cells_Rmax, size=log(maxtot_cells_Rmax), colour=as.factor(month(datemax_cells_Rmax))), alpha=0.5) + 
        ylab("Number Sampling Dates") + xlab("Sampling date") +
        scale_size_continuous("ln(Cell Dens)") +
        scale_colour_discrete("month") +
        theme(legend.position="right")

ggplot() + 
  geom_point(data=BGtotals_maxsummer, aes(y=num_sampdates_per_max, x=datemax_cells_Rmax, size=log(maxtot_cells_Rmax), colour=lake), alpha=0.5) + 
        ylab("Number Sampling Dates") + xlab("Sampling date") +
        scale_size_continuous("ln(Cell Dens)") +
        scale_colour_discrete("lake") +
        theme(legend.position="right")


ggplot() + 
  geom_point(data=BGtotals_maxsummer, aes(x=num_sampdates_per_max, y=log(maxtot_cells_Rmax), colour=as.factor(month(datemax_cells_Rmax))), alpha=0.7, size=1.8) + 
        ylab("Max ln(Total cells/L)") + xlab("Number of sample dates per year") +
        scale_colour_discrete("month") +
        theme(legend.position="top")

ggplot() + 
  geom_point(data=BGtotals_maxsummer, aes(x=num_sampdates_per_max, y=log(maxtot_cells_Rmax), colour=lake), alpha=0.7, size=1.8) + 
        ylab("Max ln(Total cells/L)") + xlab("Number of sample dates per year") +
        scale_colour_discrete("lake") +
        theme(legend.position="right")


ggplot() + 
  geom_point(data=BGtotals_maxsummer, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax), size=num_sampdates_per_max), colour="blue", alpha=0.6) + 
        ylab("max ln(Total cells/L)") + xlab("Sampling date") +
        scale_size_continuous("# Dates") +
        theme(legend.position="right")

ggplot() + 
  geom_point(data=BGtotals_maxsummer, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax), size=num_sampdates_per_max, colour=lake), alpha=0.6) + 
        ylab("max ln(Total cells/L)") + xlab("Sampling date") +
        scale_size_continuous("# Dates") +
        scale_colour_discrete("lake") +
        theme(legend.position="right")

ggplot() + 
  geom_point(data=BGtotals_maxsummer, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax), size=num_sampdates_per_max, colour=lake), alpha=0.6) + 
  geom_line(data=BGtotals_maxsummer, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax), colour=lake), alpha=0.6) +
        ylab("max ln(Total cells/L)") + xlab("Sampling date") +
        scale_size_continuous("# Dates") +
        scale_colour_discrete("lake") +
        theme(legend.position="right")

```

```{r, fig.width=7, fig.height=7}
ggplot() + 
  geom_point(data=BGtotals_maxsummer, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax), size=num_sampdates_per_max), alpha=0.6, colour="blue") + 
  geom_line(data=BGtotals_maxsummer, aes(x=datemax_cells_Rmax, y=log(maxtot_cells_Rmax)), alpha=0.6) +
   facet_wrap(~lake, nrow=5, ncol=4) +
        ylab("max ln(Total cells/L)") + xlab("Sampling date") +
        scale_size_continuous("#  Dates") +
        scale_colour_discrete("lake") +
        theme(legend.position="right")

```




# Add red points at maximums in panel plots we looked at before


```{r}
day_in_yr <- as.Date(with(BGtotals_data, paste("1000","-",substring(date1,6,10), sep="")))
BGtotals_data$day_in_yr <- day_in_yr

BGtotals_maxsummer <- mutate(BGtotals_maxsummer, day_in_yr=as.Date(paste("1000","-",substring(datemax_cells_Rmax,6,10), sep="")),
year.f=factor(year))
```

```{r, fig.width=8, fig.height=6}
for (l in lake_list) {
plot.dat <- filter(BGtotals_data, lake==l, depth_cat=="0-3 ft", !year.f=="1995")
plot.dat2 <- filter(BGtotals_maxsummer, lake==l, !year==1995)
d5 <-  ggplot(plot.dat) +
  geom_point(data=plot.dat, aes(x=day_in_yr, y=log(tot_cells_Rmax), colour=station)) + 
        geom_vline(aes(xintercept=as.numeric(as.Date("1000-06-01"))), colour="red", size=0.5) +
        geom_vline(aes(xintercept=as.numeric(as.Date("1000-10-01"))), colour="red", size=0.5) +
        geom_point(data=plot.dat2, aes(y=log(maxtot_cells_Rmax), x=day_in_yr, group=year.f), colour="red", size=3) +
        geom_hline(aes(yintercept=23.03), colour="red", size=0.5)+
        geom_hline(aes(yintercept=18.42), colour="pink", size=0.5) +
        geom_hline(aes(yintercept=16.81), colour="orange", size=0.5) +
        facet_wrap(~year.f, nrow=5, ncol=4) +
        xlab("Day of year") +
        ylab("ln(BG Total density)")+
        ggtitle(paste(l)) +
        scale_x_date(date_minor_breaks = "1 month")
print(d5)
}
```

## Lump all

```{r, fig.width=8, fig.height=6}
## Plot all points from all lakes together
for (l in lake_list) {
plot.dat <- filter(BGtotals_data, depth_cat=="0-3 ft")
plot.dat2 <- BGtotals_maxsummer

d6 <- ggplot(plot.dat) +
   geom_point(data=plot.dat, aes(x=day_in_yr, y=log(tot_cells_Rmax))) + 
        geom_vline(aes(xintercept=as.numeric(as.Date("1000-06-01"))), colour="red", size=0.5) +
        geom_vline(aes(xintercept=as.numeric(as.Date("1000-10-01"))), colour="red", size=0.5) +
        geom_point(data=plot.dat2, aes(y=log(maxtot_cells_Rmax), x=day_in_yr, group=year.f), colour="red", size=2, alpha=0.5) +
        geom_hline(aes(yintercept=23.03), colour="red", size=0.5)+
        geom_hline(aes(yintercept=18.42), colour="pink", size=0.5) +
        geom_hline(aes(yintercept=16.81), colour="orange", size=0.5) +
        facet_wrap(~year.f) +
        xlab("Day of year") +
        ylab("ln(BG Total density)")+
        ggtitle("EFR - log cell density for BG Algae") +
        scale_x_date(date_minor_breaks = "1 month")

print(d6)
}    
```





```{r CompareTrend20001toMax, eval=FALSE, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
#lake_list <- c("EFR", "CHL", "RRR", "BHR")
#l<-"EFR"
lake_list <- unique(BGtotals_data$lake)
for (l in lake_list) {
  plot.dat <- filter(BGtotals_data, lake==l, station=="20001", depth_cat=="0-3 ft")
  plot.dat2 <- filter(BGtotals_summer, lake==l, station=="20001")
  plot.dat3 <- filter(BGtotals_maxsummer, lake==l)
  #View(plot.dat2)
  
  d <- ggplot() + 
        geom_point(data=plot.dat, aes(x=date1, y=log(tot_cells_Ravg)), shape=1, size=2) +
        geom_line(data=plot.dat, aes(x=date1, y=log(tot_cells_Ravg))) + 
        geom_smooth(data=plot.dat, aes(x=date1, y=log(tot_cells_Ravg), colour="20001 all samples"), method="lm", se=FALSE) +
        geom_point(data=plot.dat2, aes(x=date1, y=log(tot_cells_Ravg), colour="20001 June-Sept"), shape=16,
                   size=1) +
        geom_line(data=plot.dat2, aes(x=date1, y=log(tot_cells_Ravg), colour="20001 June-Sept")) + 
        geom_smooth(data=plot.dat2,  aes(x=date1, y=log(tot_cells_Ravg), colour="20001 June-Sept"), method="lm", se=FALSE) +
    
        geom_point(data=plot.dat3, aes(x=datemax_cells_Ravg, y=log(maxtot_cells_Ravg), colour="Max June-Sept"), shape=10, size=2) +
        geom_line(data=plot.dat3, aes(x=datemax_cells_Ravg, y=log(maxtot_cells_Ravg)),
                    col=gray(0.3)) + 
        geom_smooth(data=plot.dat3,  aes(x=datemax_cells_Ravg, y=log(maxtot_cells_Ravg), colour="Max June-Sept"), method="lm", se=FALSE) +
    
        ggtitle(paste(l, "BG Algae - 20001 station vs. Summer Max")) +
        scale_x_date(date_minor_breaks = "1 year") +
        ylab("log(Total cells/L)") + xlab("Sampling date") +
        theme(legend.title=element_blank()) +
        theme(legend.position="top")
      print(d)
      
     # ggplot_build(d)$data
}
```
